manuscript_analysis

### 2024-07-03 manuscript_analysis.Rmd
solved problems with some of the models, checked AICs and simplified them.

df_1to3_dept_temp <- bind_rows(df_1, df_2, df_3) |> 
  filter(tot_abu > 20)

df_1to3_dept_temp |> 
  distinct(Species)

predictions_depth_temp(df_1to3_dept_temp)
# DO NOT RUN NOW, RUN LATER!
predictions_depth_temp(df_1)
predictions_depth_temp(df_2)
predictions_depth_temp(df_3)
# 4 not run yet, will wait with this
predictions_depth_temp(df_4)

stats_gamms(df_test)

@@

####Fig. 2
Overview of all model predictions

>>>>

# DO NOT RUN NOW, RUN LATER!
predictions_depth_temp(df_1)
predictions_depth_temp(df_2)
predictions_depth_temp(df_3)
predictions_depth_temp(df_4)

@@

###AIC comparisons in special cases

AIC_testing <- function(data2, i){

  data <- data2 |> 
    filter(Species == i)

  data$fProtocol <- as.factor(data$Protocol)
  data$fLake <- as.factor(data$Lake)

  model_0 <- gam(data = data, Abundance ~ s(mean_last_7days, k = 3) +
                   s(Depth_sample, k = 3)
                            + s(fLake, bs = 're')
                            +  s(fProtocol, bs = 're'), family = ziP())
  print(glance(model_0))
# new models

model1 <- gam(data = data, Presence ~ s(mean_last_7days, k = 3) + s(Depth_sample, k = 3)
                            + s(fLake, bs = 're')
                            +  s(fProtocol, bs = 're'), family = binomial)
print(glance(model1))

model2 <- gam(data = data, Abundance ~ s(mean_last_7days, k = 3) +
                            + s(fLake, bs = 're')
                            +  s(fProtocol, bs = 're'), family = ziP())
print(glance(model2))

AIC_testing(df_4, "Lepomis_gibbosus")

# in Alosa agone: model_0 not running, smaller AIC for binomial depth + temp model -> will be modelled as binomial depth + temp
# same for Coregonus_sp_large_pelagic
# lepomis gibbosus also has the smallest AIC for binomial depth + temp model

# also testing the other special cases:
 # "Alburnus_arborella", "Barbatula_sp_Lineage_I","Cyprinus_carpio", "Phoxinus_csikii", "Salmo_trutta"

AIC_testing(df_4, "Alburnus_arborella") 
# not running with model0, Binomial with depth and temp best
AIC_testing(df_4, "Barbatula_sp_Lineage_I")
# binomial depth and temp has lowest AIC
AIC_testing(df_4, "Cyprinus_carpio")
# binomial depth and temp has lowest AIC 
AIC_testing(df_4, "Phoxinus_csikii")
#binomial depth and temp has lowest AIC 
AIC_testing(df_4, "Salmo_trutta")
# binomial depth and temp lowest AIC

# 8 models are run with binomial temp and depth

```
####Fig. 2
Overview of all model predictions
